---
title: "Week 1 Quiz"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

1. In the NFL, a professional American football league, there are 32 teams, of which 12 make the playoffs. In a typical season, 20 teams (the ones that don’t make the playoffs) play 16 games, 4 teams play 17 games, 6 teams play 18 games, and 2 teams play 19 games. At the beginning of each game, a coin is flipped to determine who gets the football first. You are told that an unknown team won ten of its coin flips last season. Given this information, what is the posterior probability that the team did not make the playoffs (i.e. played 16 games)?
```{r}
p <- c(16,17,18,19)
prior<-c(20/32,4/32,6/32,2/32)
likelihood <- dbinom(10,p,prob = 1/2)
numerator <- prior * likelihood
denominator <- sum(numerator)
posterior <- numerator / denominator
posterior
```

2. Recall the probability distribution of a Poisson random variable X:
$$P(X=x)=\frac{λ^xe^{−λ}}{x!}$$
where λ is the rate parameter that equals the expected value of X.

You are a professor and assign your TAs to type up a very important homework assignment. You have three TAs: David makes an average of one typo per page, Amy makes an average of two typos per page, and Joe makes an average of three typos per page. A one-page typed homework assignment is turned into your box that has ten typos! Assuming that typos follow a Poisson distribution and you have no prior knowledge about which TA typed the assignment, what is the posterior probability that the TA who typed the homework assignment was Joe?

```{r}
p<-c(1,2,3)
prior<-c(1/3,1/3,1/3)
likelihood <- dpois(10,p)
numerator <- prior * likelihood
denominator <- sum(numerator)
posterior <- numerator / denominator
posterior
```


3. Hearing about your brilliant success in working with M&Ms, Mars Inc. transfers you over to the Skittles department. They recently have had complaints about tropical Skittles being mixed in with original Skittles. You decide to conduct a frequentist analysis. If the findings suggest that more than 1% of sampled supposedly original skittles are actually tropical, you will recommend action to be taken and the production process to be corrected. You will use a significance level of α=0.1. You randomly sample 300 supposedly original skittles, and you find that five of them are tropical. What should be the conclusion of your hypothesis test? 
Hint: H0:p=0.01 and H1:p>0.01.
```{r}
sum(dbinom(5:300,300,0.01))
```

4. You draw two balls from one of three possible large urns, labelled A, B, and C. Urn A has 1/2 blue balls, 1/3 green balls, and 1/6 red balls. Urn B has 1/6 blue balls, 1/2 green balls, and 1/3 red balls. Urn C has 1/3 blue balls, 1/6 green balls, and 1/2 red balls. With no prior information about which urn your are drawing from, you draw one red ball and one blue ball. What is the probability that you drew from urn C?

Using Bayes' Theorem, we now compute as follows: 
$$\frac{P(X|C)P(C)}{P(X|C)P(C)+P(X|A)P(A)+P(X|B)P(B)}$$
where the bracketed portion in the denominator corresponds to the cases in which red and blue were selected but did not come from Urn C.

Under the assumption that each of P(C), P(A), and P(B) is 1/3, we can cancel this factor from the numerator and denominator to get:
$$\frac{P(X|C)}{P(X|C)+P(X|A)+P(X|B)}$$
Next, we calculate each of these terms:
P(X|C)=2(3/6)(2/5)=12/30; P(X|B)=2(2/6)(1/5)=4/30; and P(X|A)=2(1/6)(3/5)=6/30.
And so the earlier expression becomes:
$$\frac{12/30}{12/30+4/30+6/30}=\frac{12}{12+4+6}=\frac{6}{6+2+3}=\frac{6}{11}$$

5. You go to Las Vegas and sit down at a slot machine. You are told by a highly reliable source that, for each spin, the probability of hitting the jackpot is either 1 in 1,000 or 1 in 1,000,000, but you have no prior information to tell you which of the two it is. You play ten times, but do not win the jackpot. What is the posterior probability that the true odds of hitting the jackpot are 1 in 1,000?
```{r}
p<-c(1/1000, 1/1000000)
prior <- c(1/2,1/2)
likelihood <- dbinom(0,10,prob = p)
numerator <- prior * likelihood
denominator <- sum(numerator)
posterior <- numerator / denominator
posterior
```

6. You decide to conduct a statistical analysis of a lottery to determine how many possible lottery combinations there were. If there are N possible lottery combinations, each person has a 1/N chance of winning. Suppose that 413,271,201 people played the lottery and three people won. You are told that the number of lottery combinations is a multiple of 100 million and less than 1 billion, but have no other prior information to go on. What is the posterior probability that there were fewer than 600 million lottery combinations?
```{r}
p <- seq(from = 100000000,to = 900000000,by = 100000000)
prior<-rep(1/9,9)
likelihood <- dbinom(3,413271201,prob = 1/p)
numerator <- prior * likelihood
denominator <- sum(numerator)
posterior <- numerator / denominator
sum(posterior[1:5])
```

7. A New York City cab was involved in a hit-and-run accident last night. Five witnesses reported the incident, four of whom said that the cab was green and one of whom said that the cab was yellow. Assume each witness correctly identifies the color of a cab with probability 2/3. It is known that 85% of registered cabs in New York City are yellow and 15% are green. Based on this information, what is the probability that the cab was green?
58.5%

8. You are testing dice for a casino to make sure that sixes do not come up more frequently than expected. Because you do not want to manually roll dice all day, you design a machine to roll a die repeatedly and record the number of sixes that come face up. In order to do a Bayesian analysis to test the hypothesis that p = 1/6 versus p = .175 , you set the machine to roll the die 6000 times. When you come back at the end of the day, you discover to your horror that the machine was unable to count higher than 999. The machine says that 999 sixes occurred. Given a prior probability of 0.8 placed on the hypothesis p = 1/6 , what is the posterior probability that the die is fair, given the censored data? Hint - to find the probability that at least x sixes occurred in N trials with proportion p (which is the likelihood in this problem), use the R command :
[1-pbinom(x-1,N,p)]

```{r}
p <- c(1/6,0.175)
prior<-c(0.8,0.2)
likelihood <- 1-pbinom(999-1,6000,p)
numerator <- prior * likelihood
denominator <- sum(numerator)
posterior <- numerator / denominator
posterior
```

9. A new breast cancer screening method is tested to see if it performs better than existing methods in detecting breast cancer. To measure the sensitivity of the test, a total of 10,000 patients known to have various stages of breast cancer are testing using the new method. Of those 10,000 patients, 9,942 are identified by the new method to have breast cancer. Given that the sensitivity of the best current test is 99.3%, is there significant evidence at the α=0.05 level to conclude that the new method has higher sensitivity than existing methods? Hint - H0:p=0.993 and H1:p>0.993
```{r}
sum(dbinom(9942:10000,10000,0.993))
```











